---
title: "Data Wrangling With R"
subtitle: "Day One"
author: "Prof. John Lalor"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: zenburn
    css: documentCSS.css
---


```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE, comment = "")
```

# Welcome to Data Wrangling with R!

## Quick Admin 

- Course Outline
- Intro Slides: any questions? 
- Introductions 
  - Name
  - Hometown 
  - Best book you've read so far in 2020
  
## Logistics for this week

This file is a knitted RMarkdown document. 
If that sentence doesn't make sense now, it will by the end of the day. 
We'll talk through a lot of R code this week, and I expect you to be following along on your laptop.

You should open a new .R file on your laptop (call it "data_wrangling_day1.R" or something like that).
As we go through code, you should type it out, run it, and see what happens. 
If you run into errors, or have questions, let me know! 
We'll figure this out together.


# The Tidyverse

```{r, echo = FALSE}
tidyverse::tidyverse_logo()
```


We are going to be working around in the <span class="pack">tidyverse</span> for a good chunk of our time together.

[Tidyverse homepage](https://www.tidyverse.org/) 

The whole point of the <span class="pack">tidyverse</span> is to offer a grammar of verbs. 
It is going to help us in a lot of the situations that we are going to be seeing.


Another great feature of the <span class="pack">tidyverse</span> is the pipe: <span class="func">%>%</span>

It does the same thing as the Unix |, but | in R is an or operator. 

With all of the glowing praise for the <span class="pack">tidyverse</span>, we are still going to see some <span class="pack">base</span> R. 
Sometimes, it will demonstrate great reasons for using the <span class="pack">tidyverse</span>. 
In other situations, it will help you to not be afraid to use it when situations arise.

## Tidyverse Note

If you need a lot of the tidyverse, you can load it.

```{r eval=FALSE}
library(tidyverse)
```
By a lot of the tidyverse I mean the core packages that are included in the tidyverse: [core packages](https://www.tidyverse.org/packages/). 
If the package you need isn't in this list, you load it directly. 
If the package you need is in this list, but it's the only one that you need, you can load it directly.
For the next few sections we'll load each package as we need it...


# Some Demonstrations 

## Summary Tables

```{r}
# load our plotting library, ggplot2
library(ggplot2)

# diamonds is a data set that comes packaged with ggplot2
plotDat = aggregate(diamonds$cut, by = list(cut = diamonds$cut), 
                    FUN = length)

# what's happening here? 
colnames(plotDat)[2] = "n"

plotDat
```

## Visual

```{r}

ggplot(plotDat, aes(x = cut, y = n)) +
  geom_point(aes(size = n)) +
  theme_minimal()
```

## (Im)Proper Plotting

Look at <span class="func">help(mtcars)</span> (also known as <span class="func">?mtcars</span>) and check out the variables. 
Can you spot what is wrong with this plot?

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = am)) + 
  geom_point() +
  theme_minimal()
```


## Proper Plotting

The plot below is likely better.

```{r}
library(dplyr)

mtcars$amFactor = as.factor(mtcars$am) 

ggplot(mtcars, aes(x = wt, y = mpg, color = amFactor)) + 
  geom_point() +
  theme_minimal()
```


## Pipes: Making Life Easier


Recall some of the things that we just saw:

```{r, eval = FALSE}
plotDat = aggregate(diamonds$cut, by = list(cut = diamonds$cut), FUN = length)

colnames(plotDat)[2] = "n"

ggplot(plotDat, aes(x = cut, y = n)) +
  geom_point(aes(size = n)) +
  theme_minimal()
```

This is somewhat tricky code. 
We have to create a new object with the oft-muddy <span class="func">aggregate</span> and reset a column name (by magic number in an index, no less). 

This can be made much easier with <span class="pack">dplyr</span>:

```{r}
# no intermediate objects between what we have (diamonds) and what we want (plot) 
diamonds %>% 
  group_by(cut) %>% 
  summarize(n = n()) %>% 
  ggplot(., aes(x = cut, y = n)) +
  geom_point(aes(size = n)) +
  theme_minimal()
  

```

It isn't a reduction in lines, but it is certainly clearer and follows a more logical thought process. 
This is the whole point of the <span class="pack">tidyverse</span> (and <span class="pack">dplyr</span> specifically) -- allowing you to write how you would explain the process. 

Remember the "grammar" I mentioned earlier?

The <span class="pack">tidyverse</span> operations each "do" something in the pipe sequence, so you can see explicitly what is being done to your data line by line. 

As an added bonus, we don't need to create a bunch of different objects to do something simple.

We can see that <span class="pack">dplyr</span> will also make the plot for am easier.

```{r}
mtcars %>% 
  mutate(am = as.factor(am)) %>%  
  ggplot(., aes(x = wt, y = mpg, color = am)) + 
  geom_point() +
  theme_minimal()
```

## On Code Golf

You will often notice that a <span class="pack">dplyr</span> chunk might take a few more lines to work through than base R alone -- don't consider this as a bad thing. 
There will be many times in this course and in your potential work that you might think that you need to use as few lines as possible. 
Resist this temptation. 
Sometime you need to break something up into many lines and create new objects -- this ability is exactly why we use R!


# Data Import 

Data import can be easy. 
Often times, it *is* easy (never too hard to import a nice .csv). 
Sometimes, though, it most certainly is not. 

## Delimited Files

Frequently, you will see nicely delimited text files that are not .csv files -- these are often tab-delimited files, but they can take other forms. 
Delimited files are the most straightforward in terms of importing, as long as you know what is the separating character.

```{r, eval = FALSE}
# will take a few seconds
read.table("https://download.bls.gov/pub/time.series/ce/ce.data.42a.RetailTrade.Employment", 
           header = TRUE, sep = "\t")
```

Is the same as:

```{r, eval = FALSE}
# will take a few seconds
read.delim("https://download.bls.gov/pub/time.series/ce/ce.data.42a.RetailTrade.Employment")
```

Where is the "sep" argument in <span class="func">read.delim()</span> above? 
Let's check the documentation...

The <span class="func">read.table()</span> function gives you added flexibility to specify many different parameters. 

### Your turn (YT1) 

Examine the following file from SDC Platinum and read it in properly:

<a href="https://www3.nd.edu/~sberry5/data/sdcTest.txt">SDC Wackiness</a>

How did you do?

### Alternatives

Did you notice anything about these files? They are not really very big, but they might have taken a little bit of time to read in. 
There have been times where people have commented that R is too slow on the read side. 
If you find you files are not being read quickly enough, you can try a few alternatives: <span class="pack">readr</span> and <span class="pack">data.table</span>

Try the following:

```{r, eval = FALSE}
# notice that we have to load the library before we can use functions in it
library(readr)

readrTest = read_delim("https://download.bls.gov/pub/time.series/ce/ce.data.42a.RetailTrade.Employment", 
                       delim = "\t")

# why are different columns colored differently in the status output?
# https://rdrr.io/github/tidyverse/readr/src/R/col_types.R 
```


```{r, eval = FALSE}
# again, load the library before we can use functions in it
library(data.table)

dtTest = fread("https://download.bls.gov/pub/time.series/ce/ce.data.42a.RetailTrade.Employment", 
               sep = "\t")
```


That SDC file that might have taken a few minutes will now take just a few seconds.
Note that read_delim (and readr more generally) tries to "guess" column specifications.
If it guesses wrong you have to provide the specs yourself (see the "col_types" argument below).
While this takes a little more work on the front end, the result is that you get data parsed the way you want it. 

```{r, eval = FALSE}
sdc = read_delim("https://www3.nd.edu/~sberry5/data/sdcTest.txt", 
                 delim="^", quote="",
                 col_types="DDDDDDcccnnnDDccccccccnl",
                 locale = locale(date_format = "%m/%d/%y"))
```


Pretty awesome, right?

While <span class="pack">readr</span> works wonderfully on the read and write side, <span class="pack">data.table</span> is great for wrangling data that is a bit on the big side and is all together blazing fast. 
However, it does not shy away from confusing syntax and weird conventions. 
With that in mind, we won't be using it in this class, but do keep it in the back of your mind.  


At times, you will get data in some proprietary format. 
That is when you need to turn to other places.

## Excel

Download the following Excel file: https://www3.nd.edu/~sberry5/data/excelTest.xlsx

What do we know about Excel workbooks? Check out the help on <span class="pack">readxl</span> and let me know our path forward.

```{r, eval = FALSE}
readxl::read_excel(path = "")
```



## SAS

```{r, eval = FALSE}
haven::read_sas(data_file = "https://www3.nd.edu/~sberry5/data/wciklink_gvkey.sas7bdat")
```


## Stata

```{r, eval = FALSE}
haven::read_dta(file = "https://www3.nd.edu/~sberry5/data/stataExample.dta")
```

## SPSS


```{r, eval = FALSE}
haven::read_spss(file = "https://www3.nd.edu/~sberry5/data/spssExample.sav", 
                 user_na = FALSE)

```
Note: We often see the -99 added as the missing value in SPSS (of course, there is no way that -99 would ever be an actual value, right?).
You can update the "user_na" argument to reflect your data.

## HTML

Depending on your needs, reading an html table into R is getting to be too easy.

```{r, eval = FALSE}
# load our library before using its packages
library(rvest)


# what is html_table doing? Let's check the help
cpi = read_html("http://www.usinflationcalculator.com/inflation/consumer-price-index-and-annual-percent-changes-from-1913-to-2008/") %>% 
  html_table(fill = TRUE)
```


Things might get a bit tricky:

```{r, eval = FALSE}
highest = read_html("https://en.wikipedia.org/wiki/List_of_highest-grossing_films") %>% 
  html_table(fill = TRUE)
```

What is the return of this call?


## rio

For many of these tasks, you can just use the <span class="pack">rio</span> package -- you give it the file and it will do the rest!

```{r, eval = FALSE}
rio::import("folder/file")
```

From rio's GitHub page [link](https://github.com/leeper/rio): 

> The core advantage of rio is that it makes assumptions that the user is probably willing to make

## Nested Structures

#### JSON

Web-based graphics started getting popular not too long ago. Generally, stats people were not using them, but web developer-type folks were. They needed a structure that would work well for the web and interact with their JavaScript-based graphics -- thus, JavaScript Object Notation (JSON) was born. You will see JSON come out of many web-based interfaces. 

<a href="https://www3.nd.edu/~sberry5/data/optionsDataBrief.json">This</a> is what JSON looks like.

There are a few JSON-reading packages in R, but jsonlite tends to work pretty well.

```{r, eval = FALSE}
jsonTest = jsonlite::read_json(path = "https://www3.nd.edu/~sberry5/data/optionsDataBrief.json", 
                                simplifyVector = TRUE)
```

A few questions:

- We've now seen it a few times, but what's with the double colon?
- What is the simplifyVector argument for?
- How do I find the answers to these questions??

This is a very simple form of JSON. 
We are going to see a hairier version of this data in the coming days. 
You may see JSON in your group project as well. 

#### JSON Dangers

There is JSON and then there is JSON. 
You might find yourself some interesting data and want to bring it in, but an error happens and you have no idea why the read_json function is telling you that the file is not JSON. 

Not all JSON is pure JSON! 
When that is the case, you will need to create pure JSON.

Look at this file: https://www3.nd.edu/~sberry5/data/reviews_Musical_Instruments_5.json

It looks like JSON, but...

```{r, eval = FALSE}
jsonlite::validate("https://www3.nd.edu/~sberry5/data/reviews_Musical_Instruments_5.json")
```

What is wrong with this file?
(Hint: let's compare it with the previous JSON file we saw)

If we would want to read that in as true JSON, we would need to do some work:

```{r, eval = FALSE}
musicalInstruments = readLines("https://www3.nd.edu/~sberry5/data/reviews_Musical_Instruments_5.json")

musicalInstruments = paste(unlist(lapply(musicalInstruments, function(x) {
  paste(x, ",", sep = "")
})), collapse = "")

musicalInstruments = paste("[", musicalInstruments, "]", sep = "")

musicalInstruments = gsub("},]", "}]", musicalInstruments)
```

What is going on in the above? (Think about what makes "pure JSON" valid...) 

## Mass Reading

Everything we just learned is great and you will use them all in your data wrangling missions.

Fortunately (or unfortunately, depending on how you look at it), it is not the whole story -- you will frequently be reading in many files of the same time.

If you have two files, you might be able to get away with brute force:

```{r, eval = FALSE}
# DO NOT RUN:

myData1 = read.csv("test.csv")

myData2 = read.csv("test2.csv")
```

Would you want to do this for 5 files? 
What about 100? Or 1000? 
I will answer it for you: no!

The chunks below introduce some very important functions. We are going to see <span class="func">lapply</span> again -- it is important that you learn to love the apply family!

```{r, eval = FALSE}
# DO NOT RUN:

allFiles = list.files(path = "", all.files = TRUE, full.names = TRUE, 
                      recursive = TRUE, include.dirs = FALSE)

# we'll talk more about lapply in a little bit...
allFilesRead = lapply(allFiles, function(x) read.csv(x, stringsAsFactors = FALSE))

allData = do.call("rbind", allFilesRead)
```


You can also use <span class="pack">rio</span>:

```{r, eval = FALSE}
# DO NOT RUN:

rio::import_list("", rbind = TRUE)
```



# The Grammar Of Data

One of the major aims of the <span class="pack">tidyverse</span> is to provide a clear and consistent grammar to data manipulation. 
This is helpful when diving deeper into the weeds. 

Do you remember this?

```{r, eval = FALSE}
highest = read_html("https://en.wikipedia.org/wiki/List_of_highest-grossing_films") %>% 
  html_table(fill = TRUE)
```

What did we get out of this? 
It was a big list of data frames. 
If we are looking for only one thing and we know that it is the first thing, we have some options:


```{r, eval = FALSE}
highest = highest[[1]]
```


This is great for keeping the object at first and then plucking out what we want. 
If you want the whole thing to be together, though, we have even more options:

```{r, eval = FALSE}
highest = read_html("https://en.wikipedia.org/wiki/List_of_highest-grossing_films") %>% 
  html_table(fill = TRUE) %>% 
  `[[`(1)
```


And now we see why R mystifies people. 
What is that bit of nonsense at the end? 
It is really just an index shortcut. 
Once you know how to use it, it is great; however, it will make you shake your head if you see it in the wild without knowing about it first.

This is where the benefit of <span class="pack">tidyverse</span> becomes clear. 


```{r, eval = FALSE}
highest = read_html("https://en.wikipedia.org/wiki/List_of_highest-grossing_films") %>% 
  html_table(fill = TRUE) %>%
  magrittr::extract2(1)
```


Or...


```{r, eval = FALSE}
highest = read_html("https://en.wikipedia.org/wiki/List_of_highest-grossing_films") %>% 
  html_table(fill = TRUE) %>%
  purrr::pluck(1)
```


Both functions are doing the same thing and with slightly different names, but it is crystal-clear what they are doing.

# Selecting 

## Base

There are many ways to select variables with <span class="pack">base</span> R:

```{r, eval = FALSE}
mtcars[, c(1:5, 7:8)]

keepers = c("mpg", "cyl", "disp", "hp", "drat", "qsec", "vs")

mtcars[, keepers]

mtcars[, c("mpg", grep("^c", names(mtcars), value = TRUE))]
```

What is going on in that last one? 
"^c" is a **regular expression** which looks for column names that begin with the letter "c".
We'll be seeing much more about regular expressions later in the week. 

You can also drop variables:

```{r}
mtcars[, -c(1:2)]

dropVars = c("vs", "drat")

mtcars[, !(names(mtcars) %in% dropVars)]
```

Issues?

For starters, the magic numbers are a no-go. 
The keepers lines could work, but would be a pain if we had a lot of variables. 

Let's check this wacky stuff out where we want all variables that start with "age" and variables that likely represent questions (x1, x2, x3, ...):

```{r}
library(lavaan)

testData = HolzingerSwineford1939

names(testData)

keepers = c(grep("^age", names(testData), value = TRUE), 
            paste("x", 1:9, sep = ""))

testData = testData[, keepers]

```


Not only do we have another regular expression, but we also have this paste line to create variable names. 
It seems like too much work to do something simple!

While not beautiful, these are perfectly valid ways to do this work. 
I have such sights to show you, but don't forget about this stuff -- you never know when you might need to use it.

## dplyr

We have already seen a bit of <span class="pack">dplyr</span>, but we are going to dive right into some of the functions now.

In base R, we have to do some chanting to select our variables. 
With <span class="pack">dplyr</span>, we can just use <span class="func">select</span>: 

```{r}
mtcars %>% 
  select(mpg, cyl, am)
```


We can also drop variables with the <span class="func">-</span>:

```{r}
mtcars %>% 
  select(-vs)
```


We also have several helper functions that we can use:

```{r, eval = FALSE}
HolzingerSwineford1939 %>% 
  select(num_range("x", 1:9), starts_with("age"), 
         matches("^s.*.l$"))
```


#### Not Important, But Helpful

Changing variable position in R is a pain:

```{r}
head(HolzingerSwineford1939[, c(1, 7:15, 2:6)])
```



```{r}
HolzingerSwineford1939 %>% 
  select(id, starts_with("x"), everything()) %>% 
  head()
```


## Your Turn! (YT2) 

1.  Use that Stata test file.

```{r, eval = FALSE}
# from above 

haven::read_dta(file = "https://www3.nd.edu/~sberry5/data/stataExample.dta")
```

2.  Grab every lvi, effect, leader, and cred variable

3.  Use <span class="func">summary</span> to understand your data.

4.  Now, just keep every lvi variable.

5.  Use a corrplot to see relationships.

    - corrplot needs a correlation matrix (use cor)

```{r, eval = FALSE}
# Just to give you an idea about how it works!

install.packages("corrplot")
library(corrplot)

data.frame(x = rnorm(10), y = rnorm(10)) %>% 
  cor() %>% 
  corrplot()
```


# Subsetting/Filtering 

One of the more frequent tasks is related to filtering/subsetting your data. 
You often want to impose some types of rules on your data (e.g., US only, date ranges).

## Base

R gives us all the ability in the world to filter data.

```{r}
summary(mtcars[mtcars$mpg < mean(mtcars$mpg), ])
```

Unless you know exactly what you are doing, this is a bit hard to read -- you might be asking yourself what the comma means and why there is nothing after it.


## dplyr

When we use <span class="func">filter</span>, we are specifying what it is that we want to keep.

Keep this or that:

```{r}
mtcars %>% 
  filter(cyl == 4 | cyl == 8) %>% 
  summary()
```

Keep this and that:

```{r}
mtcars %>% 
  filter(cyl == 4 & mpg > 25) %>% 
  summary()
```

Filter this out:

```{r}
mtcars %>% 
  filter(cyl != 4) %>% 
  summary()
```



Naturally, it can also take a function

```{r}
mtcars %>% 
  filter(mpg < mean(mpg)) %>% 
  summary()
```


## Your Turn (YT3) 

For now, we are going to stick with that stataExample data. 

1.  Select the same variables, but also include Rater.

2.  Filter the data on Rater -- check the values and filter both ways.

3.  Now check those correlations again!

4.  Throw the Gender variable in and filter on that.

# New Variables and Recoding 

## Base

Adding a new variable in <span class="pack">base</span> R is as easy as the following:

```{r}
mtcars$roundedMPG = round(mtcars$mpg)
```

## dplyr

If, however, we want to do things in a tidy chunk, we need to use <span class="func">mutate</span>.

```{r}
mtcars = mtcars %>% 
  mutate(roundedMPG = round(mpg))
```


There is also <span class="func">transmute</span>. 
Can anyone venture a guess as to what it might do?


## Base Recoding

You will need to recode variables at some point. 
Depending on the nature of the recode it can be easy (e.g., to reverse code a scale, you just subtract every value by max value + 1).

You will need to do some more elaborate stuff:

```{r, eval = FALSE}

mtcars$mpgLoHi = 0

mtcars$mpgLoHi[mtcars$mpg > median(mtcars$mpg)] = 1
```


```{r, eval = FALSE}
mtcars$mpgLoHi = ifelse(mtcars$mpg > median(mtcars$mpg), 1, 0)
```

These are pretty good ways to do recoding of this nature, but what about this:

```{r, eval = FALSE}
mtcars$vs[which(mtcars$vs == 0)] = "v"

mtcars$vs[which(mtcars$vs == 1)] = "s"
```


Or this:

```{r, eval = FALSE}
mtcars$vs = ifelse(mtcars$vs == 0, "v", "s")
```


## dplyr recoding

```{r, eval = FALSE}
recode(mtcars$vs, `0` = "v", `1` = "s")
```

## Your Turn! (YT4) 

1.  For the sake of demonstration, select only the first 10 lvi variables and everything else.

2.  Keep only observations with Rater == 0. 

3.  Assume that the first 5 lvi variables (01 through 05) are scores for one assessment and the next five (06 through 10) are scores for another assessment.

4.  Create two new variables to capture the mean of those scores.

  - You will need to use the <span class="func">rowwise</span> function ahead of mutate.
  
  - You can use the <span class="func">mean</span> function, but you will have to wrap the variables in <span class="func">c()</span>


```{r, eval = FALSE}
# Just to help you along!

data.frame(x = rnorm(10), y = rnorm(10)) %>% 
  rowwise() %>% 
  mutate(test = mean(c(x, y)))
```

# Communication

We won't have any big end-of-day wrap exercises to do today. 
Instead, we are going to learn just a few cool things.

## ggplot2

We already saw some <span class="pack">ggplot2</span>, but let's take a few minutes to dive into it a bit more.

Just like everything else in the <span class="pack">tidyverse</span>, <span class="pack">ggplot2</span> provides a clear and consistent grammar, except the focus is on data visualization. 
With <span class="pack">ggplot2</span>, we can stack layer after layer into the plotting space to help visualize our data.

Let's take a look at some good <span class="pack">ggplot2</span> layering:

```{r}
library(ggplot2)

library(lavaan)

testData = HolzingerSwineford1939

ggplot(testData, aes(x7, ageyr)) +
  geom_point()
```

Next, we can add some color:

```{r}
ggplot(testData, aes(x7, ageyr)) +
  geom_point(aes(color = as.factor(grade)), alpha = .75)
```

Now, we can add a smooth line:

```{r}
ggplot(testData, aes(x7, ageyr)) +
  geom_point(aes(color = as.factor(grade)), alpha = .75) + 
  geom_smooth()
```

And we can look at small multiples:

```{r}
ggplot(testData, aes(x7, ageyr)) +
  geom_point(aes(color = as.factor(grade)), alpha = .75) + 
  geom_smooth() +
  facet_grid(~ sex)
```

Let's get those silly grey boxes out of there:

```{r}
ggplot(testData, aes(x7, ageyr)) +
  geom_point(aes(color = as.factor(grade)), alpha = .75) + 
  geom_smooth() +
  facet_grid(~ sex) +
  theme_minimal()
```


Perhaps add a better color scheme:

```{r}
ggplot(testData, aes(x7, ageyr)) +
  geom_point(aes(color = as.factor(grade)), alpha = .75) + 
  geom_smooth() +
  facet_grid(~ sex) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2")
```


We could keep going forever and tweak anything that you could imagine (labels, ticks, etc.), but this should give you a pretty good idea about what you can do with regard to static plots.

Oh...but we don't have to stick with just static plots. 
We can use the <span class="pack">plotly</span> package to make our ggplot object interactive.

```{r}
library(plotly)

radPlot = ggplot(testData, aes(x7, ageyr)) +
  geom_point(aes(color = as.factor(grade)), alpha = .75) + 
  geom_smooth() +
  facet_grid(~ sex) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2")

ggplotly(radPlot)

```

You can also build plots with <span class="pack">plotly</span>, but we will save that for another day in the future.

Learning to use <span class="pack">ggplot2</span> will pay great dividends -- there is absolutely **nothing** better for creating visualizations. 
There is even a whole group of packages that do nothing but add stuff into it.

## DT

Visualizations are great and they often tell a better story than tables. 
Sometimes, though, you want to give people a glimpse of the data. 
The <span class="pack">DT</span> package let's you create interactive data tables (they are JS data tables). 

You could give people the entire data to explore:

```{r, eval = FALSE}
# install if necessary
# install.packages("DT")
library(DT)

datatable(testData)

```


You can also use the <span class="broom">DT</span> package to tidy your summaries into a nice data frame:

```{r, eval = FALSE}
lm(x7 ~ ageyr + school, data = testData) %>% 
  broom::tidy() %>% 
  mutate_if(is.numeric, round, 4) %>% 
  datatable()
```


We don't want to get too far ahead of ourselves here -- we will see more places to use this tomorrow.

## R Markdown & Knitr

Do you have a moment to hear the good word of Donald Knuth? 
If you want to work in a reproducible fashion <span class="rmarkdown">base</span> and <span class="pack">knitr</span> are here to help you out. 
The slides you saw earlier and even the document you are seeing now are all done with R Markdown. 
It is my hope that you will also use R Markdown for your presentations on Thursday.


