---
title: "Data Wrangling With R"
subtitle: "Introduction"
author: "Prof. John Lalor"
date: October 5, 2020
output: binb::metropolis
#  revealjs::revealjs_presentation:
#    css: documentCSS.css
#    theme: night
#    center: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Wrangling

Data wrangling is potentially comprised of many different tasks.

At the end of the day, we are trying to create data rectangles.


## Basics: Quickly

Everything in R is an object -- everything. single values, models, and even functions -- all objects.

To create an object, we just assign it a name.

```{r}
x = 1:5

y <- 6:10
```

You can use either "=" or "<-".

- The shortcut for "<-" is Alt + -
  - In Rmd files, Alt + - gives you "="

## Assignment

The stray puppy:

```{r}
1:5
```

Naming the stray:

```{r}
fido = 1:5
```

Grooming the stray:

```{r}
fido = fido * 2
```

## 

```{r}
fido
```


## Applying Functions

Now that we have some objects, x and y, we can apply some functions to each:


```{r}
xMean = mean(x)

ySD = sd(y)

print(c(xMean, ySD)) 
```

- why "c(xMean, ySD)" ?

## Combining Functions

We can also combine functions!

```{r}

abs(mean(-1:-10))

```

Clear as mud, right?

## Packages

Many functions are grouped into things called packages. 

Packages are usually written to contain functions that do similar things.

You load a package like this:

```{r, eval = FALSE}
library(dplyr)
```


If you don't have a package already, you install it with:

```{r, eval = FALSE}
install.packages("dplyr")
```

## Important Issues

All closures need...closing: (), {}, '', ""

Case matters:

  -  thisName != ThisName != this_name
  
## Our Packages

These are the packages that you will need:

```{r, eval = FALSE}
install.packages(c("tidyverse", "rio", 
                   "jsonlite", "psych", 
                   "lavaan", "reshape", 
                   "reshape2", "stringdist"))
```

If more creep up, we will install them when the time is right.

## R and RStudio

Download and Install R

- [http://cran.us.r-project.org/](http://cran.us.r-project.org/)

Download and Install RStudio

- Integrated Development Environment (IDE) for R
- [https://rstudio.com/products/rstudio/download/](https://rstudio.com/products/rstudio/download/)

## Keep this in mind...

- The R/Rstudio community is awesome

- R (and in particular the tidyverse) is *very* well documented
  - If you forget something, it's online
  - If you don't know how to do something, it's online
- Resources (a few years old but still relevant):
  - The R community is one of R's best features
    - [https://blog.revolutionanalytics.com/2017/06/r-community.html](https://blog.revolutionanalytics.com/2017/06/r-community.html)
  - Hey! You there! You are welcome here
    - [https://ropensci.org/blog/2017/06/23/community/](https://ropensci.org/blog/2017/06/23/community/)
  
  
